APPM report - Draft

### 2022-12-19 paper_version/main.tex
roman's comments and a new version of abstract

A key coefficient characterizing plasmas is the so-called Debye length $\lambda_D$, whose magnitude indicates to what extent net electric charges deviate from zero. It may range over many orders of magnitude: If $\lambda_D = \mathcal{O}(1)$ we face the non-neutral regime, while for $\lambda_D\to 0$ the plasma becomes quasi-neutral. A singularly perturbed problem arises from the $\lambda_D\to 0$ limit, which leads to peculiar issues in numerical algorithms when both regimes coexist, for instance, in the fringes of electric arcs and near electrodes. It is desired to design numerical schemes that are \emph{robust} for arbitrary $\lambda_D$. More precisely, they should be \emph{asymptotic-preserving} (AP), in the sense that the limit $\lambda_D\to 0$ of the scheme yields a viable discretization for the continuous limit model.

An asymptotic-preserving scheme for single- and multi-fluid Euler-Maxwell system was proposed by \cite{degond_2012} for one spatial dimension. We extend their scheme to three dimensions. The key ingredients are (i) a discretization of Maxwell's equations based on primal and dual meshes in the spirit of discrete exterior calculus (DEC)/the finite integration technique (FIT), (ii) finite volume method (FVM) for the fluid equations on the dual mesh, and (iii) mixed implicit-explicit timestepping. The scheme is designed to be AP as $\lambda_D \rightarrow 0$ and this is validated through numerical experiments. Besides, if the electromagnetic fields have to be modelled in an non-conducting region beyond the plasma domain, additional stabilization to accommodate Gauss's law is necessary.

@@

\mycomment{Roman: more info from a physical point of view?}

We consider a hydrodynamic multi-species plasma model, which boils down to the Maxwell-Euler system. As in \cite{degond_2012}, our focus is an \emph{asymptotic-preserving} (AP) discretization in the quasi-neutral limit $\lambda_D \rightarrow 0$, where $\lambda_D$ is the rescaled Debye length, see Section \ref{sec:rescaling} for the underlying scaling arguments and \cite{degond_2012,degond_2017} for a discussion of the importance of the AP property in plasma codes.

@@

For the sake of simplicity, we only consider ion-electron collision in this work, but extension to more species is straightforward. Such collisions can be well modelled by friction terms (\citep[][sec. 5.6.2]{chen2016})
\begin{equation} \label{equ:collision}
    \mathbf{R}_{e}^{\text{coll}} = \frac{\pi Ze^4m_e^{1/2}}{(4\pi\epsilon)^2(k_BT_e)^{3/2}}\text{ln}\Lambda n_en_i(\mathbf{u}_i - \mathbf{u}_e) = - \mathbf{R}_{i}^{\text{coll}},  
\end{equation}
which is the collision force acting on electrons. ln$\Lambda$ is called the \emph{Coulomb logarithm} and is usually taken as a constant \cite{chen2016}. Because of the conservation of momentum, the collision term for ions $\mathbf{R}_{i}^{\text{coll}}$ is nothing but the negative of $\mathbf{R}_{e}^{\text{coll}}$. In this work, the other effects, e.g. ionization and recombination, are not considered. For full details about the modeling of these physical processes, readers are referred to \cite[][Sec. 3.2]{fuchs_2021}.

@@

\subsection{Quasi-neutral Limit} \label{sec:quasi-neutral_limit}
The Debye length $\lambda_D$ (see its definition in the preceding section) is the characteristic length over which local concentrations of charge are shielded, see \cite[sec. 1.4]{chen2016} for the underlying physics. In many situations, it has a rather low value compared to the reference length scale: the typical $\lambda_D$ for gas discharge and Tokamak is $10^{-4}$ meter; for interstellar medium, the magnitude is about $1$ meter\cite[ch. 20]{Blandford_2013}. That justifies the quasi-neutrality assumption that the plasma is locally neutral, i.e. $\rho = \sum_* q_*n_* = 0$. The limit model is obtained by formally setting the rescaled Debye length $\lambda$ to zero:
\vspace{-\baselineskip}
\begin{center}
    \begin{minipage}{0.3\textwidth}
    \begin{align*}
        \partial_t \mathbf{B} + \nabla \times \mathbf{E} &= 0 \tag{\ref{equ:maxwell_faraday_rescaled}},\\ 
        \lambda^2 \partial_t \mathbf{D} - \nabla \times \mathbf{H} &= - \mathbf{J}, \tag{\ref{equ:maxwell_ampere_rescaled}}\\
        \nabla \cdot \mathbf{B} &= 0, \tag{\ref{equ:maxwell_gauss_B_rescaled}}\\
        \lambda^2 \nabla \cdot \mathbf{D} &= \rho. \tag{\ref{equ:maxwell_gauss_D_rescaled}}
    \end{align*}
\end{minipage}
\hspace{0.6cm}$\xrightarrow[]{\lambda \rightarrow 0}$
\begin{minipage}{0.3\textwidth}
\begin{subequations}
\begin{align}
    \partial_t \mathbf{B} + \nabla \times \mathbf{E} &= 0, \label{equ:maxwell_faraday_limit} \\ 
    - \nabla \times \mathbf{H} &= - \mathbf{J}, \label{equ:maxwell_ampere_limit} \\
    \nabla \cdot \mathbf{B} &= 0,  \label{equ:maxwell_gauss_B_limit}\\
     0 &= \rho. \label{equ:maxwell_gauss_D_limit}
\end{align}
\end{subequations}
\end{minipage}
\end{center}
The quasi-neutral condition (\ref{equ:maxwell_gauss_D_limit}) is the consequence of setting $\lambda = 0$ in Gauss's law for the electric field (\ref{equ:maxwell_gauss_D_rescaled}). As in the non-neutral case, Gauss's laws (\ref{equ:maxwell_gauss_B_limit}) (\ref{equ:maxwell_gauss_D_limit}) are the consequence of (\ref{equ:maxwell_faraday_limit}) (\ref{equ:maxwell_ampere_limit}) provided (\ref{equ:maxwell_gauss_B_limit}) (\ref{equ:maxwell_gauss_D_limit}) are satisfied at initial time.

@@

\begin{remark}
    The preceeding formal limit argument is heuristic. Deep mathematical tools from perturbation theory would be needed to rigorously prove the convergence of the limiting sequence. This topic is beyond the scope of this paper. Readers are referred to \cite[ch. 2]{remi_2014} \cite{Peng_2008} for the quasi-neutral limit of the Maxwell-Euler system, and \cite{mark_1995, Eckhaus_1980} for the general theory.
\end{remark}

@@

We consider a simply-connected domain $\Omega$ comprising two sub-domains $\Omega_P$ (plamsa domain) and $\Omega_N$ (non-conducting domain). They represent regions of space occupied by plasma and a non-conducting linear dielectric medium (e.g. glass) respectively. We use the notation $\Gamma_C := \partial\Omega \cap \partial\Omega_P$ to denote the contact boundary, and $\Gamma_I := \partial\Omega \cap \partial\Omega_N$ for the insulating boundary with an artificial cut-off boundary. The interface between the two domains is denoted by $\Gamma_W = \partial\Omega_I \cap \partial\Omega_P$. A geometric arrangement corresponding to this setting is sketched in Fig. \ref{fig:domain_sketch} and is the one used in \cite[][Ch. 1, sec. 4]{fuchs_2021}.

\mycomment{Roman: expand on a relevant physical setting, e.g. wall-stabilized arc in hight voltage circuit breakers.}

The Euler system is defined in $\Omega_P$ with boundary $\partial\Omega_P$. We consider the physical scenario where plasma is confined by a solid insulator. Therefore, the interface $\Gamma_W$ is treated as reflective wall where fluid particles are bounced back while keeping the tangential velocity and energy unchanged. The contact boundary $\Gamma_C$ is where the plasma and energy flow across, modelled by open boundaries in the sense that fluids are free to flow through, i.e. ions and electrons are emitted and absorbed by the electrodes.

@@

\todo{Reference to github repository}

@@

Our approach has been tested on a cylindrical domain with ECE boundary conditions and the AP property has been verified numerically. To the best of the authors' knowledge, this work is among the first to try to solve the 3D fluid plasma model on unstructured meshes, which potentially gives more flexibility than the finite-difference-based approaches and reduces the computational cost compared to particle-based approaches. In particular, the AP property allows us to address regimes of the arbitrary value of the Debye length.

\todo{Polish the reference}
\bibliography{ref}
\end{document}

>>>>

\noindent\textbf{A short version:}

The Debye length $\lambda_D$, as a key characteristic coefficient of plasmas, may range over many orders of magnitude. The induced singularly perturbed problem is a challenge for plasma simulations. Recently so called asymptotic-preserving (AP) schemes, which target such problems, are studied widely. However, in the context of plasma simulations, only AP schemes in one spatial dimension are available. In this work, we construct an AP schemes for Maxwell-Euler equations in three spatial dimensions based on primal-dual meshes and the finite integration technique (FIT) for Maxwell's equations coupled with the finite volume method (FVM) for the fluid equations. Besides, issues like boundary treatment and stabilization in non-conducting regions are tackled. The AP property is verified numerically. We expect this new discretization approach is particularly suited for problems involving multiple scales of the Debye length, e.g. electric arcs.

\noindent\textbf{A long version:}

{\color{red}
A key coefficient characterizing plasmas is the so-called Debye length $\lambda_D$, whose magnitude indicates to what extent net electric charges deviate from zero. It may range over many orders of magnitude: If $\lambda_D = \mathcal{O}(1)$ we face the non-neutral regime, while for $\lambda_D\to 0$ the plasma becomes quasi-neutral. A singularly perturbed problem arises from the $\lambda_D\to 0$ limit, which leads to peculiar issues in numerical algorithms when both regimes coexist, for instance, in the fringes of electric arcs and near electrodes. It is desired to design numerical schemes that are \emph{robust} for arbitrary $\lambda_D$. More precisely, they should be \emph{asymptotic-preserving} (AP), in the sense that the limit $\lambda_D\to 0$ of the scheme yields a viable discretization for the continuous limit model.

An asymptotic-preserving scheme for single- and multi-fluid Euler-Maxwell system was proposed by \cite{degond_2012} for one spatial dimension. We extend their scheme to three dimensions. The key ingredients are (i) a discretization of Maxwell's equations based on primal and dual meshes in the spirit of discrete exterior calculus (DEC)/the finite integration technique (FIT), (ii) finite volume method (FVM) for the fluid equations on the dual mesh, and (iii) mixed implicit-explicit timestepping. The scheme is designed to be AP as $\lambda_D \rightarrow 0$ and this is validated through numerical experiments. Besides, if the electromagnetic fields have to be modelled in an non-conducting region beyond the plasma domain, additional stabilization to accommodate Gauss's law is necessary.
}

@@

% \mycomment{Roman: more info from a physical point of view?} R.H. 19.12. Extended paragraph
The starting point for this work was the desire to numerically simulate the formation and evolution of electric arcs at atmospheric pressures. We rely on a widely accepted description by means of a hydrodynamic multi-species plasma model, which boils down to the Maxwell-Euler system. The arc phenomenon covers a wide range of plasma regimes and the numerical model should be able to deal with all of them seamlessly and simultaneously. Thus in this work, as in \cite{degond_2012}, the focus is on \emph{asymptotic-preserving} (AP) discretization in the quasi-neutral limit $\lambda \rightarrow 0$, where $\lambda$ is the rescaled Debye length, see Section \ref{sec:rescaling} for the underlying scaling arguments and \cite{degond_2012,degond_2017} for a discussion of the importance of the AP property of numerical plasma models.

@@

For the sake of simplicity, we only consider ion-electron collision in this work, but extension to more species is straightforward. Such collisions can be well modelled by friction terms (\citep[][Sec. 5.6.2]{chen2016})
\begin{equation} \label{equ:collision}
    \mathbf{R}_{e}^{\text{coll}} = \frac{\pi Ze^4m_e^{1/2}}{(4\pi\epsilon)^2(k_BT_e)^{3/2}}\text{ln}\Lambda n_en_i(\mathbf{u}_i - \mathbf{u}_e) = - \mathbf{R}_{i}^{\text{coll}},  
\end{equation}
which is the collision force acting on electrons. ln$\Lambda$ is called the \emph{Coulomb logarithm} and is usually taken as a constant \cite{chen2016}. Because of the conservation of momentum, the collision term for ions $\mathbf{R}_{i}^{\text{coll}}$ is nothing but the negative of $\mathbf{R}_{e}^{\text{coll}}$. In this work, the other effects, e.g. ionization and recombination, are not considered. For full details about the modeling of these physical processes, readers are referred to \cite[][Sec. 3.2]{fuchs_2021}.

@@

\subsection{Quasi-neutral Limit} \label{sec:quasi-neutral_limit}
The Debye length $\lambda_D$ (see its definition in the preceding section) is the characteristic length over which local concentrations of charge are shielded, see \cite[Sec. 1.4]{chen2016} for the underlying physics. In many situations, it has a rather low value compared to the reference length scale: the typical $\lambda_D$ for gas discharge and Tokamak is $10^{-4}$ meter; for interstellar medium, the magnitude is about $1$ meter\cite[Ch. 20]{Blandford_2013}. That justifies the quasi-neutrality assumption that the plasma is locally neutral, i.e. $\rho = \sum_* q_*n_* = 0$. The limit model is obtained by formally setting the rescaled Debye length $\lambda$ to zero:
\vspace{-\baselineskip}
\begin{center}
    \begin{minipage}{0.3\textwidth}
    \begin{align*}
        \partial_t \mathbf{B} + \nabla \times \mathbf{E} &= 0 \tag{\ref{equ:maxwell_faraday_rescaled}},\\ 
        \lambda^2 \partial_t \mathbf{D} - \nabla \times \mathbf{H} &= - \mathbf{J}, \tag{\ref{equ:maxwell_ampere_rescaled}}\\
        \nabla \cdot \mathbf{B} &= 0, \tag{\ref{equ:maxwell_gauss_B_rescaled}}\\
        \lambda^2 \nabla \cdot \mathbf{D} &= \rho. \tag{\ref{equ:maxwell_gauss_D_rescaled}}
    \end{align*}
\end{minipage}
\hspace{0.6cm}$\xrightarrow[]{\lambda \rightarrow 0}$
\begin{minipage}{0.3\textwidth}
\begin{subequations}
\begin{align}
    \partial_t \mathbf{B} + \nabla \times \mathbf{E} &= 0, \label{equ:maxwell_faraday_limit} \\ 
    - \nabla \times \mathbf{H} &= - \mathbf{J}, \label{equ:maxwell_ampere_limit} \\
    \nabla \cdot \mathbf{B} &= 0,  \label{equ:maxwell_gauss_B_limit}\\
     0 &= \rho. \label{equ:maxwell_gauss_D_limit}
\end{align}
\end{subequations}
\end{minipage}
\end{center}
The quasi-neutral condition (\ref{equ:maxwell_gauss_D_limit}) is the consequence of setting $\lambda = 0$ in Gauss's law for the electric field (\ref{equ:maxwell_gauss_D_rescaled}). As in the non-neutral case, Gauss's laws (\ref{equ:maxwell_gauss_B_limit}) (\ref{equ:maxwell_gauss_D_limit}) are the consequence of (\ref{equ:maxwell_faraday_limit}) (\ref{equ:maxwell_ampere_limit}) provided (\ref{equ:maxwell_gauss_B_limit}) (\ref{equ:maxwell_gauss_D_limit}) are satisfied at initial time.

@@

\begin{remark}
    The preceeding formal limit argument is heuristic. Deep mathematical tools from perturbation theory would be needed to rigorously prove the convergence of the limiting sequence. This topic is beyond the scope of this paper. Readers are referred to \cite[Ch. 2]{remi_2014} \cite{Peng_2008} for the quasi-neutral limit of the Maxwell-Euler system, and \cite{mark_1995, Eckhaus_1980} for the general theory.
\end{remark}

@@

We consider a simply-connected domain $\Omega$ comprising two sub-domains $\Omega_P$ (plamsa domain) and $\Omega_N$ (non-conducting domain). They represent regions of space occupied by plasma and a non-conducting linear dielectric medium (e.g. glass) respectively. We use the notation $\Gamma_C := \partial\Omega \cap \partial\Omega_P$ to denote the contact boundary, and $\Gamma_I := \partial\Omega \cap \partial\Omega_N$ for the insulating boundary with an artificial cut-off boundary. The interface between the two domains is denoted by $\Gamma_W = \partial\Omega_I \cap \partial\Omega_P$. A geometric arrangement corresponding to this setting is sketched in Fig. \ref{fig:domain_sketch} and is the one used in \cite[][Ch. 1, Sec. 4]{fuchs_2021}. This situation is simple, but highly relevant. It covers fundamental experimental set-ups, for example, those studying wall-stabilized electric arcs \cite{GGF05}, but is also adequate for a range of industrial plasmas, for instance, those occurring in high-voltage circuit breakers \cite{SEE15,MGU20}. 
%\mycomment{Roman: expand on a relevant physical setting, e.g. wall-stabilized arc in hight voltage circuit breakers.}

The Euler equations are posed $\Omega_P$ and associated boundary conditions are imposed on $\partial\Omega_P$. We consider the physical scenario where plasma is confined by a solid insulator. Therefore, the interface $\Gamma_W$ is treated as reflective wall where fluid particles are bounced back while keeping the tangential velocity and energy unchanged. The contact boundary $\Gamma_C$ is where the plasma and energy flow across, modelled by open boundaries in the sense that fluids are free to flow through, i.e. ions and electrons are emitted and absorbed by the electrodes.

@@

@@

Our approach has been tested on a cylindrical domain with settings meant for electric arc simulations. The AP property has been verified numerically. To the best of the authors' knowledge, this work is among the first to try to solve the 3D fluid plasma model on unstructured meshes, which potentially gives more flexibility than the finite-difference-based approaches and reduces the computational cost compared to particle-based approaches. In particular, the AP property allows us to address regimes of the arbitrary value of the Debye length.

\bibliography{ref}
\end{document}
